HUME: Human UCCA-Based Evaluation of Machine Translation

### 2016-03-18 paper/main.tex
shortened section 1

Human judgement is the central, and perhaps most fundamental, criterion for
estimating the quality of an MT system.
Nevertheless, common measures for MT evaluation, such as adequacy and fluency judgements
or the relative ranking of several possible translations, are problematic in two ways.
First, as the quality of translation is determined by multiple factors, it is difficult
to assign a single number that quantifies the quality of the entire sentence. This
is indeed reflected in the diminishing inter-annotator agreement (IAA) rates of human ranking measures
%as a function of
with
the sentence length \cite{Bojar:2011}.
Second, a sentence-level quality score does not indicate what parts of the sentence
are erroneously translated, and hence cannot inform developers in overcoming these errors.

These problems are partially addressed by measures that decompose over sub-parts of the evaluated
translation sentence (henceforth, {\it translation}). For automatic measures,
these are commonly words or n-grams, for manual measures some structural
information is taken into account \parcite{machacek:bojar:segranks:2015}.
% the following is specific to automatic measures:
%quantifying the overlap of its sub-parts relative a reference translation.
A promising line of research decomposes
%More recent work on MT evaluation proposed measures
%that decompose
over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure; the most notable of these measures is
HMEANT \cite{lo2011structured}.
\XXX{Do we want to cut the following shorter?}
These measures can localize semantic errors in the translation, but
their focus on verb argument structures misses many pervasive semantic phenomena
(such as inter-clause linkage and nominal argument structures) and thus they only provide
a limited perspective on the semantic similarity between the output and reference.
Moreover, assigning semantic structure to MT output, especially when low in quality,
is difficult and arguably ill-defined.

We propose the Human UCCA-based MT Evaluation (HUME) metric,
a human evaluation measure that decomposes over the semantic units of the sentence.
Semantic units are defined according to the 
UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
due to its cross-linguistic applicability, support for rapid annotation, and coverage
of many fundamental semantic phenomena, such as verbal, nominal and adjectival
argument structures and the inter-relations between them.

HUME operates by aggregating human assessments of the translation quality of individual
semantic units in the source sentence.
Importantly, HUME only requires semantically annotating the source sentence,
thus avoiding the annotation of machine-generated text 
and allowing re-use of the source semantic annotation for measuring the quality
of different translations of the same source sentence.
By basing the measure on a direct comparison
of the source and translation we avoid the difficulties engendered by using
reference translations, albeit at the cost of imposing the additional requirement
that the evaluator be proficient in both the source and target languages.
%An elaborate comparison of HUME and HMEANT is given in \secref{sec:hmeant_comp}.

After a brief review (\secref{sec:background}), we describe HUME in detail
(\secref{sec:hume}) and highlight its benefits over HMEANT
(\secref{sec:hmeant_comp}).
Our experiments with four language pairs: English-German, English-Polish,
English-Romanian and English-Czech (\secref{sec:experiments}) document the
efficiency (time of annotation) and inter-annotator agreement scores for HUME.

@@

\section{Conclusion}

>>>>

Human judgement is the most important criterion for
estimating the quality of an MT system.
Nevertheless, common measures for human MT evaluation, such as adequacy and fluency judgements
or the relative ranking of possible translations, are problematic in two ways.
First, as the quality of translation is determined by multiple factors, it is difficult
to assign a single number that quantifies the quality of the entire sentence. This
is indeed reflected in the diminishing inter-annotator agreement (IAA) rates of human ranking measures
%as a function of
with
the sentence length \cite{Bojar:2011}.
Second, a sentence-level quality score does not indicate what parts of the sentence
are badly translated, and so cannot inform developers in repairing these errors.

These problems are partially addressed by measures that decompose over parts of the evaluated
translation. For automatic measures,
these are commonly words or n-grams, for manual measures some structural
information is taken into account \parcite{machacek:bojar:segranks:2015}.
% the following is specific to automatic measures:
%quantifying the overlap of its sub-parts relative a reference translation.
A promising line of research decomposes
%More recent work on MT evaluation proposed measures
%that decompose
over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure; the most notable of these measures is
HMEANT \cite{lo2011structured}.
\XXX{Do we want to cut the following shorter?}
However, 
their focus on verb argument structures misses many semantic phenomena
(such as inter-clause linkage and nominal argument structures) and thus they only provide
a limited perspective on the semantic similarity between the translation and reference.
Also, assigning semantic structure to an MT output, especially a poor one,
is difficult and arguably ill-defined.

We propose the Human UCCA-based MT Evaluation (HUME) metric,
a human evaluation measure that decomposes over the semantic units of the sentence.
Semantic units are defined according to the 
UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
due to its cross-linguistic applicability, support for rapid annotation, and coverage
of many fundamental semantic phenomena, such as verbal, nominal and adjectival
argument structures and their inter-relations.

HUME operates by aggregating human assessments of the translation quality of individual
semantic units in the source sentence.
HUME only requires semantically annotating the source sentence,
thus avoiding the annotation of machine-generated text 
and allowing re-use of the source semantic annotation for measuring the quality
of different translations of the same source sentence.
By basing the measure on a direct comparison
of the source and translation we avoid the difficulties created by using
reference translations, albeit at the cost of requiring the
evaluator to be proficient in both the source and target languages.
%An elaborate comparison of HUME and HMEANT is given in \secref{sec:hmeant_comp}.

After a brief review (\secref{sec:background}), we describe HUME in detail
(\secref{sec:hume}) and highlight its benefits over HMEANT
(\secref{sec:hmeant_comp}).
Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments}) document the
efficiency (time of annotation) and inter-annotator agreement scores for HUME.

@@
